Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-10 ICML-submission/from-linear-separability-to-strong-linear-separability.tex
Add lemma bounding the norm of linear functions

In this section, we show how to construct a mapping $\phi$ from the unit ball of
$\R^d$ into a high dimensional inner product space that has the property that if
a set of labeled examples in the unit ball is linearly separable with a margin
$\gamma$, applying the mapping $\phi$ makes the examples \emph{strongly}
linearly separable with a margin $\gamma'$ and their norms are bounded by $R'$.
The parameters $\gamma'$ and $R'$ are functions of the old margin $\gamma$ and
the number of classes $K$, and the are specified in the theorems below.

@@

Before we state the theorems, recall that a polynomial of $d$ variables is a
function $p:\R^d \to \R$ of the form
$$
p(x) = p(x_1, x_2, \dots, x_d) = \sum_{\alpha_1, \alpha_2, \dots, \alpha_d} c_{\alpha_1, \alpha_2, \dots, \alpha_d} x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d}
$$
where the sum ranges over a finite set of $d$-tuples $(\alpha_1, \alpha_2,
\dots, \alpha_d)$ of non-negative integers and $c_{\alpha_1, \alpha_2, \dots,
\alpha_d}$ is a real coefficient. The \emph{degree} of a polynomial $p$, denoted
by $\deg(p)$, is the largest value of $\alpha_1 + \alpha_2 + \dots + \alpha_d$
for which the coefficient $c_{\alpha_1, \alpha_2, \dots, \alpha_d}$ is non-zero.
The \emph{norm of a polynomial} $p$ is defined as
$$
\norm{p} = \sqrt{\sum_{\alpha_1, \alpha_2, \dots, \alpha_d} \left(c_{\alpha_1, \alpha_2, \dots, \alpha_d} \right)^2 } \; .
$$
It is easy see that this is indeed a norm, since we can interpret it as a
Euclidean norm of the vector of the coefficients of the polynomial.

@@

\begin{theorem}[Polynomial approximation of intersection of halfspaces II]
\label{theorem:polynomial-approximation-2}
Let $v_1, v_2, \dots, v_m \in \R^d$ be vectors such that $\norm{v_1},
\norm{v_2}, \dots, \norm{v_m} \le 1$. Let $\gamma \in (0,1)$.
Define
$$
r = 2 \left\lceil \frac{1}{4} \log_2(4m + 1) \right\rceil + 1 \qquad \text{and} \qquad s = \left \lceil \log_2(1/\gamma) \right \rceil \; .
$$
Then, there exists a multivariate polynomial $p:\R^d \to \R$ such that
\begin{enumerate}
\item $\displaystyle p(x) \ge \frac{1}{4} \cdot 2^{s(s+1)rm}$ \\
for all $\displaystyle x \in \bigcap_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \ge \gamma \right\}$

@@

We construct the three different mappings $\phi_n, \psi_n, \rho$. The first two
are parameterized by a non-negative integer $n$. The domain of $\phi_n$ and
$\psi_n$ is $\R^d$. The domain of $\rho$ is the unit ball of $\R^d$. The
co-domain for $\phi_n$ and $\psi_n$ is $\R^p$ where $p = \sum_{i=0}^n
\binom{i+d-1}{d-1}$ equipped with the standard inner product $\ip{x}{x'} =
\sum_{i=1}^p x_i x'_i$. The co-domain is the classical separable Hilbert space
$\ell_2 = \{ x \in \R^\infty ~:~ \sum_{i=1}^\infty x_i^2 < + \infty \}$ equipped
with the standard inner product $\ip{x}{x'} = \sum_{i=1}^\infty x_i x'_i$. We
index the coordinates of the co-domain $\R^p$ by $d$-tuples $(\alpha_1,
\alpha_2, \dots, \alpha_d)$ of non-negative integers such that $\alpha_1 +
\alpha_2 + \dots + \alpha_d \le n$. Similarly, we index the coordinates of
$\ell_2$ by $d$-tuples $(\alpha_1, \alpha_2, \dots, \alpha_d)$ of non-negative
integers. The mappings are defined by
\begin{align*}
& \left(\phi_n(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)}
= x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d}
\\
& \left(\psi_n(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)} = x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d} \\
& \qquad \cdot \sqrt{\binom{n}{\alpha_1 + \alpha_2 + \dots + \alpha_d} \binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}
\\
& \left(\rho(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)} = x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d} \\
& \qquad \cdot \sqrt{2^{-(\alpha_1 + \alpha_2 + \dots + \alpha_d)} \binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}
\end{align*}
where $\binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d} = \frac{(\alpha_1 + \alpha_2 + \dots + \alpha_d)!}{\alpha_1! \alpha_2! \dots \alpha_d!}$
is the multinomial coefficient and $\binom{n}{k} = \frac{n!}{(n-k)!k!}$ is the binomial coefficient.

Mappings $\psi_n$ and $\rho$ have simple-to-compute kernels
\begin{align*}
K_{\psi_n}(x,x') & = \ip{\psi_n(x)}{\psi_n(x')}_{\R^p} = \left( 1 + \ip{x}{x'} \right)^n \; , \\
K_{\rho}(x,x') & = \ip{\rho(x)}{\rho(x')}_{\ell_2} = \frac{1}{1 - \frac{1}{2}\ip{x}{x'}} \; .
\end{align*}
In passing we note that the formula for $K_\rho$ implies the mapping $\rho$ is
well defined. Indeed, since $K_{\rho}(x,x) < +\infty$ for any $x$ in the unit
ball of $\R^d$, it means that $\rho(x)$ indeed lies in $\ell_2$.

TODO

>>>>

In this section, we show how to construct a mapping $\phi$ from the unit ball of
$\R^d$ into a high dimensional inner product space that has the property that if
a set of labeled examples in the unit ball is linearly separable with a margin
$\gamma$, applying the mapping $\phi$ makes the examples \emph{strongly}
linearly separable with a margin $\gamma'$ and their norms are bounded by $R'$.
The parameters $\gamma'$ and $R'$ are functions of the old margin $\gamma$ and
the number of classes $K$, and are specified in the theorems below.

@@

Before we state the theorems, recall that a polynomial of $d$ variables is a
function $p:\R^d \to \R$ of the form
\begin{align*}
p(x)
& = p(x_1, x_2, \dots, x_d) \\
& = \sum_{\alpha_1, \alpha_2, \dots, \alpha_d} c_{\alpha_1, \alpha_2, \dots, \alpha_d} x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d}
\end{align*}
where the sum ranges over a finite set of $d$-tuples $(\alpha_1, \alpha_2,
\dots, \alpha_d)$ of non-negative integers and $c_{\alpha_1, \alpha_2, \dots,
\alpha_d}$ is a real coefficient. The \emph{degree} of a polynomial $p$, denoted
by $\deg(p)$, is the largest value of $\alpha_1 + \alpha_2 + \dots + \alpha_d$
for which the coefficient $c_{\alpha_1, \alpha_2, \dots, \alpha_d}$ is non-zero.
The \emph{norm of a polynomial} $p$ is defined as
$$
\norm{p} = \sqrt{\sum_{\alpha_1, \alpha_2, \dots, \alpha_d} \left(c_{\alpha_1, \alpha_2, \dots, \alpha_d} \right)^2 } \; .
$$
It is easy see that this is indeed a norm, since we can interpret it as a
Euclidean norm of the vector of the coefficients of the polynomial.

@@

\begin{theorem}[Polynomial approximation of intersection of halfspaces II]
\label{theorem:polynomial-approximation-2}
Let $v_1, v_2, \dots, v_m \in \R^d$ be vectors such that $\norm{v_1},
\norm{v_2}, \dots, \norm{v_m} \le 1$. Let $\gamma \in (0,1)$.
Define
$$
r = 2 \left\lceil \frac{1}{4} \log_2(4m + 1) \right\rceil + 1 \quad \text{and} \quad s = \left \lceil \log_2(1/\gamma) \right \rceil \; .
$$
Then, there exists a multivariate polynomial $p:\R^d \to \R$ such that
\begin{enumerate}
\item $\displaystyle p(x) \ge \frac{1}{4} \cdot 2^{s(s+1)rm}$ \\
for all $\displaystyle x \in \bigcap_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \ge \gamma \right\}$

@@

We construct the three different mappings $\phi_n, \psi_n, \rho$. The first two
are parameterized by a non-negative integer $n$. The domain of $\phi_n$ and
$\psi_n$ is $\R^d$. The domain of $\rho$ is the unit ball of $\R^d$. The
co-domain of $\phi_n$ and $\psi_n$ is $\R^q$ where $q = \sum_{i=0}^n
\binom{i+d-1}{d-1}$ equipped with the standard inner product $\ip{x}{x'}_{\R_q} =
\sum_{i=1}^p x_i x'_i$. The co-domain of $\rho$ is the classical real separable Hilbert space
$\ell_2 = \{ x \in \R^\infty ~:~ \sum_{i=1}^\infty x_i^2 < + \infty \}$ equipped
with the standard inner product $\ip{x}{x'}_{\ell_2} = \sum_{i=1}^\infty x_i x'_i$. We
index the coordinates of the co-domain $\R^q$ by $d$-tuples $(\alpha_1,
\alpha_2, \dots, \alpha_d)$ of non-negative integers such that $\alpha_1 +
\alpha_2 + \dots + \alpha_d \le n$. Similarly, we index the coordinates of
$\ell_2$ by $d$-tuples $(\alpha_1, \alpha_2, \dots, \alpha_d)$ of non-negative
integers. The mappings are defined by
\begin{align*}
& \left(\phi_n(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)}
= x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d}
\\
& \left(\psi_n(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)} = x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d} \\
& \qquad \cdot \sqrt{\binom{n}{\alpha_1 + \alpha_2 + \dots + \alpha_d} \binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}
\\
& \left(\rho(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)} = x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d} \\
& \qquad \cdot \sqrt{2^{-(\alpha_1 + \alpha_2 + \dots + \alpha_d)} \binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}
\end{align*}
where $\binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d} = \frac{(\alpha_1 + \alpha_2 + \dots + \alpha_d)!}{\alpha_1! \alpha_2! \dots \alpha_d!}$
is the multinomial coefficient and $\binom{n}{k} = \frac{n!}{(n-k)!k!}$ is the binomial coefficient.

Mappings $\psi_n$ and $\rho$ have simple-to-compute kernels
\begin{align*}
K_{\psi_n}(x,x') & = \ip{\psi_n(x)}{\psi_n(x')}_{\R^q} = \left( 1 + \ip{x}{x'} \right)^n \; , \\
K_{\rho}(x,x') & = \ip{\rho(x)}{\rho(x')}_{\ell_2} = \frac{1}{1 - \frac{1}{2}\ip{x}{x'}} \; .
\end{align*}
In passing we note that the formula for $K_\rho$ implies the mapping $\rho$ is
well defined. Indeed, since $K_{\rho}(x,x) < +\infty$ for any $x$ in the unit
ball of $\R^d$, it means that $\rho(x)$ indeed lies in $\ell_2$.

Any mutlivariate polynomial in the orginal space $\R^d$ can be expressed as a
linear function in the new feature spaces given by mappings $\phi_n, \psi_n,
\rho$. The following lemma bounds the norm of the linear function in the terms
of the norm of the polynomial. Recall that a linear function in a Hilbert space
$(H, \ip{\cdot}{\cdot}_H)$ has the form $\ip{c}{x}_H$ for some vector $c \in H$
and its norm is $\norm{c}_H = \sqrt{\ip{c}{c}}$.

\begin{lemma}
Let $d \ge 1$ and $n \ge 0$ be integers and let $q = \sum_{i=0}^n \binom{i+d-1}{d-1}$.
Let $p:\R^d \to \R$ be a multivariate polynomial.
\begin{enumerate}
\item If $n \ge \deg(p)$ then there exists $c \in \R^q$ such that $p(x) = \ip{c}{\phi_n(x)}_{\R^q}$
and $\norm{c}_{\R^q} = \norm{p}$.
\item If $n \ge \deg(p)$ then there exists $c' \in \R^q$ such that $p(x) = \ip{c'}{\psi_n(x)}_{\R^q}$
and $\norm{c'}_{\R^q} \le \norm{p}$.
\item There exists $c'' \in \ell_2$ such that $p(x) = \ip{c''}{\psi_n(x)}_{\ell_2}$
and $\norm{c''}_{\R^q} \le 2^{\deg(p)/2} \norm{p}$.
\end{enumerate}
\end{lemma}

\begin{proof}
Note that the polynomial $p$ can be written as
$p(x) = \sum_{\alpha_1, \alpha_2, \dots, \alpha_d} c_{\alpha_1, \alpha_2, \dots, \alpha_d} x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d}$.
There are $\sum_{i=0}^{\deg(p)} \binom{i+d-1}{d-1}$ coefficients $c_{\alpha_1, \alpha_2, \dots, \alpha_d}$.
If $n \ge \deg(p)$ these can be expressed as a vector in $c \in \R^q$. Thus, $p(x) = \ip{c}{\phi_n(x)}_{\R^q}$
and
$$
\norm{c}_{\R^p} = \sqrt{\sum_{\alpha_1, \alpha_2, \dots, \alpha_d} (c_{\alpha_1, \alpha_2, \dots, \alpha_d})^2} =  \norm{p} \; .
$$

We define $c' \in \R^q$ using the multi-index notation as
$$
c'_{\alpha_1, \alpha_2, \dots, \alpha_d}
= \frac{c_{\alpha_1, \alpha_2, \dots, \alpha_d}}{\sqrt{\binom{n}{\alpha_1 + \alpha_2 + \dots + \alpha_d} \binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}} \; .
$$
We have $\ip{c'}{\psi_n(x)}_{\R^q} = \ip{c}{\phi_n(x)}_{\R^q} = p(x)$. Since
$$
|c'_{\alpha_1, \alpha_2, \dots, \alpha_d}| \le |c_{\alpha_1, \alpha_2, \dots, \alpha_d}| \; ,
$$
we have $\norm{c'}_{\R^q} \le \norm{c}_{\R^q} = \norm{p}$.

Similarly, we define $c'' \in \ell_2$ using the multi-index notation as
$$
c''_{\alpha_1, \alpha_2, \dots, \alpha_d}
= \frac{c_{\alpha_1, \alpha_2, \dots, \alpha_d} 2^{(\alpha_1 + \alpha_2 + \dots + \alpha_d)/2}}{\sqrt{\binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}}
$$
for all tuples $(\alpha_1, \alpha_2, \dots, \alpha_d)$ such that $\alpha_1 + \alpha_2 + \dots + \alpha_d \le \deg(p)$.
Otherwise, we define $c''_{\alpha_1, \alpha_2, \dots, \alpha_d} = 0$. We have
$\ip{c''}{\rho(x)}_{\ell_2} = \ip{c}{\phi_n(x)}_{\R^q} = p(x)$. Since
\begin{align*}
|c''_{\alpha_1, \alpha_2, \dots, \alpha_d}|
& \le 2^{(\alpha_1 + \alpha_2 + \dots + \alpha_d)/2} |c''_{\alpha_1, \alpha_2, \dots, \alpha_d}| \\
& \le 2^{\deg(p)/2} |c''_{\alpha_1, \alpha_2, \dots, \alpha_d}| \; ,
\end{align*}
we have $\norm{c''}_{\ell_2} \le 2^{\deg(p)/2} \norm{c}_{\R^q} = 2^{\deg(p)/2} \norm{p}$.
\end{proof}
